---
title: "Programaming Assignment 3"
date: "Last updated `r Sys.time()`"
author: "Isabelle Chang"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, 'setup', message=FALSE, echo=FALSE}
#message=false: code is still echoed, but message is gone
library(tidyverse)
library(languageR)
```
## Tidy data
```{r, 'load and tidy english dataset', out.width="100%", message=FALSE, echo=FALSE}
tidy_english <- english %>%
  filter(AgeSubject=="young" & NounFrequency != 0 & VerbFrequency != 0) %>%
  #filter(NounFrequency < 500 & VerbFrequency < 500) %>%
  select(Word, NounFrequency, VerbFrequency, WordCategory) %>%
  arrange(NounFrequency, VerbFrequency, Word)

tidy_english
```

## Descriptive Statistics
```{r, 'tidy stats', echo=FALSE}
tidy_english %>%
  summarize(noun_avg=mean(NounFrequency), noun_sd=sd(NounFrequency))

tidy_english %>%  
  summarize(verb_avg=mean(VerbFrequency), verb_sd=sd(VerbFrequency))
```

## Plot:

```{r, 'plot with outliers', out.width="100%", message=FALSE, echo=FALSE}
mod1 <- lm(VerbFrequency ~ NounFrequency, data = tidy_english)

plot1 <- tidy_english %>%
  ggplot() +
  aes(x=NounFrequency, y=VerbFrequency) +
  geom_point() +
  geom_smooth(method=lm, formula="y~x")

plot1
```

### To account for outliers, I removed data points with noun or verb frequencies above 15000. The descriptive statistics and plot are as follows:
```{r, 'stats and plot without outliers', out.width='100%', message=FALSE, echo=FALSE}

tidier_english <- tidy_english %>%
  filter(VerbFrequency < 15000 & NounFrequency < 15000)

tidier_english

tidier_english %>%
  summarize(verb_avg=mean(VerbFrequency), verb_sd=sd(VerbFrequency))

mod2 <- lm(VerbFrequency ~ NounFrequency, data = tidier_english)

plot2 <- tidier_english %>%
  ggplot() +
  aes(x=NounFrequency, y=VerbFrequency) +
  geom_point() +
  geom_smooth(method=lm, formula="y~x")

plot2
```

## Analysis

The plots suggest that the relationship between a word's verb and noun frequencies is nonlinear. Data points fan out from the calculated line of best fit rather than congregating in a blob-like form around the line.